\pkg{regsem}: Regularized Structural Equation Modeling

### 2017-03-08 draft1.tex
update sim table
draft sent to sarfaraz

\Abstract{
The \pkg{regsem} package in \proglang{R}, an implementation of
Regularized Structural Equation Modeling (Jacobucci, Grimm, and McArdle
2016), was recently developed with the goal incorporating various forms
of penalized likelihood estimation with a broad array of structural
equations models. The forms of regularization include both the
\textit{ridge} (Hoerl and Kennard 1970) and the least absolute shrinkage
and selection operator (\textit{lasso}; (Tibshirani 1996), along with
sparser extensions. The paper details both the algorithmic details and
an overview of the use of \pkg{regsem} through both a factor analysis
and latent growth curve models.
}

@@

\Address{
    Ross Jacobucci\\
  University of Notre Dame\\
  First line Second line\\
  E-mail: \href{mailto:rcjacobuc@gmail.com}{\nolinkurl{rcjacobuc@gmail.com}}\\
  URL: rjacobucci.com\\~\\
  }

@@

The desire for simplicity in model structure comes by many names,
including simple structure (Thurstone 1935), variable complexity (Browne
2001), parsimony (Raykov and Marcoulides 1999; Marsh and Hau 1996),
``sparse loadings'' in the context of principal components analysis
(Zou, Hastie, and Tibshirani 2006), and lastly, ``sparsistency'',
denoting that all parameters in a sparse model that are zero are
correctly estimated as zero with probability tending to one (Lam and Fan
2009). The goal is accurately and efficiently estimate a model that is
parsimonious in allowing users to easily interpret the model's
representation of reality. In Figure 1, this would entail testing
whether the cross-loadings are necessary to the fit of the model, or
whether they can be pared away without an meaningful increase in bias.

In the context of latent variables, reducing the complexity of models
can come in many forms: selecting among multiple predictors of a latent
variable, simplifying factor structure by removing cross-loadings,
determining whether the addition of nonlinear terms are necessary in
longitudinal models, and many others. As a simple running example,
Figure 1 depicts a linear latent growth curve model with four time
points and ten predictors for a simulated dataset.

@@

As an example, a researcher may want to test this model, but may only
have a relatively small sample size (e.g.~80). However, there are 29
estimated parameters in this model, resulting in a estimated parameter
to sample size ratio as far below even the most liberal recommendations
(e.g.~10:1 parameters to sample size; Kline 2015). In lieu of finding
additional respondents, reducing the number of parameters estimated is
one effective strategy. Specifically, the 20 estimated regressions from
\textit{c1-c10} could be reduced to a number that makes the number of
parameters estimated to sample size ratio more reasonable.

@@

\[\tag{1}
\hat{\beta}^{ridge}= argmin \Big\{ \sum_{i=1}^{N} (y_{i} = \beta_{0} - \sum_{j=1}^{p}x_{ij} \beta_{j})^{2}  + \lambda \sum_{j=1}^{p} \beta_{j}^{2}\Big\},
\] \noindent
where \(\beta_{0}\) is the intercept, \(\beta_{j}\) is the coefficient
for \(x_{j}\), and \(\lambda\) is the penalty that controls the amount
of shrinkage. Note that when \(\lambda = 0\), equation 3 reduces to
ordinary least squares regression. As \(\lambda\) is increased, the
\(\beta\) parameters are shrunken towards zero. The lasso estimates are
defined as

@@

In the context of our example depicted in Figure 1, to use lasso
regression to select among the covariates, the growth model would need
to be reduced to two factor scores, which neglects both the relationship
between both the slope and intercept, reducing both to independent
factor scores. Particularly in models with a greater number of latent
variables, this becomes increasing problematic. A method that keeps the
model structure, while allowing for a penalized estimation of specific
parameters is regularized structural equation modeling (RegSEM;
Jacobucci, Grimm, and McArdle 2016). RegSEM adds a penalty function to
the traditional maximum likelihood cost function for SEMs. The maximum
likelihood cost function for SEMs can be written as

@@

\noindent
where \(\Sigma\) is the model implied covariance matrix, \(C\) is the
observed covariance matrix, and \(p\) is the number of estimated
parameters. RegSem builds in an additional element to penalize certain
model parameters yielding

@@

\noindent
where \(\lambda\) is the regularization parameter and takes on a value
between zero and infinity. When \(\lambda\) is zero, MLE is performed,
and when \(\lambda\) is infinity, all penalized parameters are shrunk to
zero. \(P(\cdot)\) is a general function for summing the values of one
or more of the model's parameter matrices. The two common forms of
\(P(\cdot)\) include both the Lasso (\(\| \cdot \|_{1}\)), which
penalizes the sum of the absolute values of the parameters, and Ridge
(\(\| \cdot \|_{2}\)), which penalizes the sum of the squared values of
the parameters.

@@

\[\tag{5}
F_{lasso} = F_{ML} + \lambda * \left\|  \begin{matrix}  
c1 \xrightarrow[]{} i\\
c2\xrightarrow[]{}i\\
\vdots \\
c10\xrightarrow[]{}i\\
c1\xrightarrow[]{}s\\
\vdots \\
c10\xrightarrow[]{}s\\
\end{matrix}  \right\|_{1}
\]

Although the fit of the model is easily calculated given a set of
parameter estimates, traditional optimization procedures for SEM cannot
be used given the non-differentiable nature of lasso penalties, and as
detailed later, extensions.

@@

One method that has become popular for optimizing penalized likelihood
method is that of proximal gradient descent (e.g.~p.~104 in Hastie,
Tibshirani, and Wainwright 2015). In comparison to one-step procedures
common in SEM optimization, that only involve a method for calculating
the step size and the direction (typically using the gradient and an
approximation of the Hessian), proximal gradient descent can be
formulated as a two-step procedure. With a stepsize of \(\alpha^{t}\)
and parameters \(\theta^{t}\) at iteration \textit{t}:

\begin{enumerate}
    \item First, take a gradient step size $z = \theta^{t} - \alpha^{t} \nabla g(\theta^{t})$.
    \item Second, perform elementwise soft-thresholding $\theta^{t+1} = S_{\alpha^{t} \lambda}(z)$.
\end{enumerate}

where \(S_{\alpha \lambda}(z)\) is the soft-thresholding operator
(Donoho 1995) used to overcome non-differentiability of the lasso
penalty at the origin:

\begin{equation}
S_{\alpha \lambda}(z_{j}) = sign(z_{j})(|z_{j}|-\alpha \lambda)_{+}
\end{equation}

where \((x)_{+}\) is shorthand for max(x,0) and \(\alpha\) is the step
size. Henceforth, \(\lambda\) is assumed to encompass both the penalty
and the step size \(s^{t}\). This procedure is only used to updated
parameters that are subject to penalty. Non-penalized parameters are
updated only using step 1 from above.

@@

\begin{minipage}{\textwidth}
    %   \renewcommand\footnoterule{}    
    \begin{algorithm}[H]
        \begin{algorithmic}[1]
            %   \Procedure{RegSEM Block Coordinate Descent for Lasso}{}
            %   \footnotetext{Note: $A(pen)$ refers only to the penalized parameters, $vec()$ concatenates both vectors}
            \State Generate starting values for $\theta_{t}$
            \State Calculate initial fit $F_{t}$
            \State Set step size $\alpha$. .5 works well at this time.
            \State Set tolerance (tol). e.g. 1e-6
            \While {$|F_{t} - F_{t+1}| > tol$}
            \State Calculate gradient for A:$ \nabla (A) =:  \frac{\partial A}{\partial \theta_{t}}$
            \State $ \theta_{t+1^{*},A} =: \theta_{t,A} - \alpha \nabla(A)$
            \State  Update penalized parameters: $\theta_{t+1^{*},A(pen)} =: S_{\alpha \lambda}(\theta_{t+1^{*},A(pen)})$
            \State Calculate gradient for S:$ \nabla (S) =:  \frac{\partial S}{\partial \theta_{t+1^{*}}}$ 
            \State Update S parameters: $ \theta_{t+1^{*},S} =: \theta_{t,S} - \alpha \nabla(S)$
            \State $\theta_{t+1} =: \theta_{t+1^{*}}$ %vec(\theta_{t+1^{*},A},\theta_{t+1^{*},S})$
            \State Update $S_{t+1}, A_{t+1}$
            \State $ \Sigma_{t+1} = F(I-A_{t+1})^{-1}S_{t+1}(I-A_{t+1})^{-T}F^{T}$
            \State $ F_{t+1} = F_{ML}(\Sigma_{t+1},C) + \lambda \| A_{t+1}(pen) \|$
            \EndWhile
            %   \EndProcedure
        \end{algorithmic}
        \caption{RegSEM Block Coordinate Descent}
        \label{alg:seq}
    \end{algorithm}
\end{minipage}

@@

Most notably, the elastic net (Zou and Hastie 2005) encompasses both the
ridge and lasso, reaching a compromise between both through the addition
of an additional parameter \(\alpha\).

@@

In using lasso penalties, difficulties emerge when the scale of
variables differ dramatically. By only using one value of \(\lambda\),
this can add appreciable bias to the resulting estimates (e.g. Fan and
Li 2001). One method proposed for overcoming this limitation is the
adaptive lasso (Zou 2006). Instead of penalizing parameters directly,
each parameter is scaled by the un-penalized estimated (maximum
likelihood estimate in SEM). This manifests itself as: \[
F_{alasso} = F_{ML} + \lambda \| \theta_{ML}^{-1} * \theta_{pen} \|_{1}
\] \noindent
with, following the same form for the lasso, the soft-thresholding
update is:

\[
S(\theta_{j})= sign(\theta_{j})(|\theta_{j}|-\frac{\lambda}{2|\theta_{j}|})_{+}
\] \noindent
In this, larger penalties are given for non-significant (smaller)
parameters, limiting the bias in estimating larger, significant,
parameters. Note that one limitation of this approach for SEM models is
that the model needs to be estimable with maximum likelihood.
Particularly for models with large numbers of variables, in relation to
sample size, this may not be possible.

@@

Two additional penalties that overcome some of the deficiencies of the
lasso, producing sparser solutions, include the smoothly clipped
absolute deviation penalty (SCAD; Fan and Li 2001) and the minimax
concave penalty (MCP; Zhang 2010). In comparison to the lasso, both the
SCAD and MCP have much steeper penalties for smaller parameters, as is
evident in Figure **.

@@

\noindent
As the the penalty in equation 11 is non-convex (as is the MCP), this
makes computation more difficult. However, in the context of SEM this
can be seen as less problematic, as equation 3 is also non-convex.

@@

\noindent
As seen in Figure **, this results in similar amount of shrinkage for
smaller estimates in comparison to the SCAD, however, less for larger
estimates.

@@

RegSEM is implemented as the \pkg{regsem} package in the \proglang{R}
statistical environment. To estimate the maximum likelihood fit of the
model, \pkg{regsem} uses \textit{Reticular Action Model} (RAM; J. J.
McArdle and McDonald 1984, McArdle (2005)) notation to derive an implied
covariance matrix. The parameters of each SEM are translated into three
matrices: the \textit{filter} (\textit{F}), the \textit{asymmetric}
(\textit{A}; directed paths; e.g.~factor loadings or regressions), and
the \textit{symmetric} (\textit{S}; undirected paths; e.g.~covariances
or variances). See Jacobucci, Grimm, and McArdle (2016) for more detail
on RAM notation.

Syntax for using the \pkg{regsem} is based on the \pkg{lavaan} package
(Rosseel 2012) for structural equation models. \pkg{lavaan} is a general
SEM software program that can fit a wide array of models with various
estimation methods. To use \pkg{regsem}, the user has to first fit the
model in \pkg{lavaan}. As an example, below is the code for a one factor
model fit in lavaan with the Holzinger-Swineford (Holzinger and
Swineford 1939) dataset.

@@

The \code{out.reg} object contains two components, \code{out.reg[[1]]}
has the parameter estimates for each of the 15 models, while
\code{out.reg[[2]]} has the fit of each model. For the code above, using
the \pkg{xtable} package to turn the output from \code{out.reg[[2]]}
into a table

\begin{CodeChunk}
\begin{CodeInput}
round(out.reg[[2]],3)
\end{CodeInput}
\begin{CodeOutput}
      lambda conv rmsea      BIC
 [1,]   0.00    0 0.188 7805.177
 [2,]   0.05    0 0.191 7815.389
 [3,]   0.10    0 0.199 7840.590
 [4,]   0.15    0 0.210 7875.842
 [5,]   0.20    0 0.203 7886.147
 [6,]   0.25    0 0.210 7911.816
 [7,]   0.30    0 0.214 7934.454
 [8,]   0.35    0 0.220 7960.670
 [9,]   0.40    0 0.227 7991.033
[10,]   0.45    0 0.229 8009.416
[11,]   0.50    0 0.234 8034.805
[12,]   0.55    0 0.242 8069.487
[13,]   0.60    0 0.286 8360.539
[14,]   0.65    0 0.286 8360.530
[15,]   0.70    0 0.286 8360.525
\end{CodeOutput}
\end{CodeChunk}

In this, the user can examine the penalty (lambda), whether the model
converged (``conv''), with values of 0 indicating convergence, and the
fit of each model. By default, two fit indices are output, both the root
mean square error of approximation (RMSEA; Steiger and Lind 1980), and
the Bayesian information criteria (BIC; Schwarz 1978). Both the RMSEA
and BIC take into account the degrees of freedom of the model, an
important point for model selection in the presence of lasso penalties
(and other penalties that set parameters to zero). This concurs with
work done on degrees of freedom for lasso regression, as Zou, Hastie,
and Tibshirani (2007) proved that the number of nonzero coefficients is
an unbiased estimate of the degrees of freedom for regression. As the
penalty increases, select parameter are set to zero, thus the degrees of
freedom increases, which for fit indices that include the degrees of
freedom in the calculation, means that although the likelihood of the
model may only get worse (increase), both the RMSEA and BIC can improve
(decrease) due to the impact of increased degrees of freedom.

Instead of examining the \code{out.reg[[1]]} output matrix of parameter
estimates, it is easiest to plot the trajectory of each of the penalized
parameters. This is accomplished with \code{plot_cv(out.reg,pars=1:9)},
resulting in:

@@

\begin{center}\includegraphics{draft1_files/figure-latex/unnamed-chunk-5-1} \end{center}

@@

After a final model (penalty) is chosen, user's have the option either
just use the output from \code{cv_regsem()}, or the final model can be
re-run with either \code{regsem()} or \code{multi_optim()} to attain
additional information. In the model above, the best fitting penalty,
according to the BIC, is \(\lambda=0\). However, for demonstration
purposes, we can choose a penalty of 0.2. With this, the model is re-run
with \code{regsem()}

@@

For the purposes of comparison, each of the 20 covariate regressions
were penalized using the lasso, adaptive lasso, SCAD, and MCP, and
compared to the maximum likelihood estimates (see Appendix for code). In
this model, the data were simulated to have two large effects (both
\textit{c1} parameters), two small effects (both \textit{c2} parameters)
and sixteen true zero effects (\textit{c3-c10} parameters). Note that
the covariates were simulated to have zero covariance among each
variable. If there was substantial collinearity among covariates, the
elastic net would be more appropriate to simultaneously select
predictors while also accounting for the collinearity. The parameter
estimates corresponding the the best fit of the BIC are displayed in
Table *

@@

While every regularization method erroneously set both simulated true
intercept effects as zero (as in maximum likelihood, which accounts for
these errors), both the adaptive lasso and SCAD correctly identified
every true zero effect. The lasso identified two false effects while the
MCP mistakenly identified one. Additionally, the lasso estimation of the
true effects was attentuated in comparison to the other regularization
methods. This is in line with previous research (cite), necessitating
the use of a two-step relaxed lasso method (Meinshausen 2007, see
Jacobucci, Grimm, and McArdle (2016)) As expected given the small ratio
between number of estimated parameters and sample size, maximum
likelihood mistakenly identified 3 false effects as significant.

To compare the performance of each penalization method further,
particularly in the presence of small number of parameters to sample
size ratios, a small scale simulation study was conducted. The same
model and effects was kept, but the sample size was varied to include
80, 200, and 1000 to demonstrate how maximum likelihood estimation of
effects improves as sample size increases, while each of the
regularization methods performs well regardless of sample size. Each run
was replicated 200 times. The results are displayed in Table *

>>>>

\Abstract{
The \pkg{regsem} package in \proglang{R}, an implementation of
regularized structural equation modeling (RegSEM; Jacobucci, Grimm, and
McArdle 2016), was recently developed with the goal incorporating
various forms of penalized likelihood estimation with a broad array of
structural equations models. The forms of regularization include both
the \textit{ridge} (Hoerl and Kennard 1970) and the least absolute
shrinkage and selection operator (\textit{lasso}; Tibshirani 1996),
along with sparser extensions. RegSEM is particularly useful for
structural equation models that have a small parameter to sample size
ratio, as the addition of penalties can reduce the complexity, thus
reducing the bias of the parameter estimates. The paper details the
algorithmic details and an overview of the use of \pkg{regsem} with the
application of both factor analysis and latent growth curve models.
}

@@

\Address{
    Ross Jacobucci\\
  University of Notre Dame\\
  118 Haggar Hall, Notre Dame, IN 46556\\
  E-mail: \href{mailto:rcjacobuc@gmail.com}{\nolinkurl{rcjacobuc@gmail.com}}\\
  URL: rjacobucci.com\\~\\
  }

@@

The desire for simplicity in model structure comes by many names,
including simple structure (Thurstone 1935), variable complexity (Browne
2001), parsimony (Raykov and Marcoulides 1999; Marsh and Hau 1996),
``sparse loadings'' in the context of principal components analysis
(Zou, Hastie, and Tibshirani 2006), and lastly, ``sparsistency'',
denoting that all parameters in a sparse model that are zero are
correctly estimated as zero with probability tending to one (Lam and Fan
2009). The goal is accurately and efficiently estimate a model that is
parsimonious in allowing users to easily interpret the model's
representation of reality. In the context of latent variables, reducing
the complexity of models can come in many forms: selecting among
multiple predictors of a latent variable, simplifying factor structure
by removing cross-loadings, determining whether the addition of
nonlinear terms are necessary in longitudinal models, and many others.
As a simple running example, Figure 1 depicts a linear latent growth
curve model (e.g. Meredith and Tisak 1990) with four time points and ten
predictors for a simulated dataset.

@@

As an example, a researcher may want to test this model, but may only
have a relatively small sample size (e.g.~80). However, there are 29
estimated parameters in this model, resulting in a estimated parameter
to sample size ratio as far below even the most liberal recommendations
(e.g.~10:1 parameters to sample size; Kline 2015). In lieu of finding
additional respondents, reducing the number of parameters estimated is
one effective strategy for reducing bias. Specifically, the 20 estimated
regressions from \textit{c1-c10} could be reduced to a number that makes
the number of parameters estimated to sample size ratio more reasonable.
To explore this further, the next section provides an overview of
regularization, and how different forms can be used to perform variable
selection across a broad range of models.

@@

\[\tag{1}
\hat{\beta}^{ridge}= argmin \Big\{ \sum_{i=1}^{N} (y_{i} = \beta_{0} - \sum_{j=1}^{p}x_{ij} \beta_{j})^{2}  + \lambda \sum_{j=1}^{p} \beta_{j}^{2}\Big\},
\] \noindent
where \(\beta_{0}\) is the intercept, \(\beta_{j}\) is the coefficient
for \(x_{j}\), and \(\lambda\) is the penalty that controls the amount
of shrinkage. Note that when \(\lambda = 0\), Equation 3 reduces to
ordinary least squares regression. As \(\lambda\) is increased, the
\(\beta\) parameters are shrunken towards zero. The lasso estimates are
defined as

@@

In the context of our example depicted in Figure 1, to use lasso
regression to select among the covariates, the growth model would need
to be reduced to two factor scores, which neglects both the relationship
between both the slope and intercept, reducing both to independent
factor scores. Particularly in models with a greater number of latent
variables, this becomes increasingly problematic. A method that keeps
the model structure, while allowing for penalized estimation of specific
parameters is regularized structural equation modeling (RegSEM;
Jacobucci, Grimm, and McArdle 2016). RegSEM adds a penalty function to
the traditional maximum likelihood estimation (MLE) for structural
equation models (SEMs). The maximum likelihood cost function for SEMs
can be written as

@@

\noindent
where \(\Sigma\) is the model implied covariance matrix, \(C\) is the
observed covariance matrix, and \(p\) is the number of estimated
parameters. RegSEM builds in an additional element to penalize certain
model parameters yielding

@@

\noindent
where \(\lambda\) is the regularization parameter and takes on a value
between zero and infinity. When \(\lambda\) is zero, MLE is performed,
and when \(\lambda\) is infinity, all penalized parameters are shrunk to
zero. \(P(\cdot)\) is a general function for summing the values of one
or more of the model's parameter matrices. The two common forms of
\(P(\cdot)\) include both the lasso (\(\| \cdot \|_{1}\)), which
penalizes the sum of the absolute values of the parameters, and ridge
(\(\| \cdot \|_{2}\)), which penalizes the sum of the squared values of
the parameters.

@@

\[\tag{5}
F_{lasso} = F_{ML} + \lambda * \left\|  \begin{matrix}  
c1 \xrightarrow[]{} i\\
c2\xrightarrow[]{}i\\
\vdots \\
c10\xrightarrow[]{}i\\
c1\xrightarrow[]{}s\\
\vdots \\
c10\xrightarrow[]{}s\\
\end{matrix}  \right\|_{1}
\] \noindent
Although the fit of the model is easily calculated given a set of
parameter estimates, traditional optimization procedures for SEM cannot
be used given the non-differentiable nature of lasso penalties, and as
detailed later, sparse extensions.

@@

One method that has become popular for optimizing penalized likelihood
method is that of proximal gradient descent (e.g.~p.~104 in Hastie,
Tibshirani, and Wainwright 2015). In comparison to one-step procedures
common in SEM optimization, that only involve a method for calculating
the step size and the direction (typically using the gradient and an
approximation of the Hessian), proximal gradient descent can be
formulated as a two-step procedure. With a stepsize of \(s^{t}\) and
parameters \(\theta^{t}\) at iteration \textit{t}:

\begin{enumerate}
    \item First, take a gradient step size $z = \theta^{t} - s^{t} \nabla g(\theta^{t})$.
    \item Second, perform elementwise soft-thresholding $\theta^{t+1} = S_{s^{t} \lambda}(z)$.
\end{enumerate}

where \(S_{s^{t} \lambda}(z)\) is the soft-thresholding operator (Donoho
1995) used to overcome non-differentiability of the lasso penalty at the
origin:

\begin{equation}
S_{s^{t} \lambda}(z_{j}) = sign(z_{j})(|z_{j}|-s^{t} \lambda)_{+}
\end{equation}

\noindent
where \((x)_{+}\) is shorthand for max(x,0) and \(s^{t}\) is the step
size. Henceforth, \(\lambda\) is assumed to encompass both the penalty
and the step size \(s^{t}\). This procedure is only used to updated
parameters that are subject to penalty. Non-penalized parameters are
updated only using step 1 from above.

@@

\begin{minipage}{\textwidth}
    %   \renewcommand\footnoterule{}    
    \begin{algorithm}[H]
        \begin{algorithmic}[1]
            %   \Procedure{RegSEM Block Coordinate Descent for Lasso}{}
            %   \footnotetext{Note: $A(pen)$ refers only to the penalized parameters, $vec()$ concatenates both vectors}
            \State Generate starting values for $\theta_{t}$
            \State Calculate initial fit $F_{t}$
            \State Set step size $s$. .1 works well at this time.
            \State Set tolerance (tol). e.g. 1e-6
            \While {$|F_{t} - F_{t+1}| > tol$}
            \State Calculate gradient for A:$ \nabla (A) =:  \frac{\partial A}{\partial \theta_{t}}$
            \State $ \theta_{t+1^{*},A} =: \theta_{t,A} - s \nabla(A)$
            \State  Update penalized parameters: $\theta_{t+1^{*},A(pen)} =: S_{s \lambda}(\theta_{t+1^{*},A(pen)})$
            \State Calculate gradient for S:$ \nabla (S) =:  \frac{\partial S}{\partial \theta_{t+1^{*}}}$ 
            \State Update S parameters: $ \theta_{t+1^{*},S} =: \theta_{t,S} - s \nabla(S)$
            \State $\theta_{t+1} =: \theta_{t+1^{*}}$ %vec(\theta_{t+1^{*},A},\theta_{t+1^{*},S})$
            \State Update $S_{t+1}, A_{t+1}$
            \State $ \Sigma_{t+1} = F(I-A_{t+1})^{-1}S_{t+1}(I-A_{t+1})^{-T}F^{T}$
            \State $ F_{t+1} = F_{ML}(\Sigma_{t+1},C) + \lambda \| A_{t+1}(pen) \|$
            \EndWhile
            %   \EndProcedure
        \end{algorithmic}
        \caption{RegSEM Block Coordinate Descent}
        \label{alg:seq}
    \end{algorithm}
\end{minipage}

@@

Most notably, the elastic net (Zou and Hastie 2005) encompasses both the
ridge and lasso, reaching a compromise between both through the addition
of an additional parameter \(\alpha\), manifesting itself as

@@

In using lasso penalties, difficulties emerge when the scale of
variables differ dramatically. By only using one value of \(\lambda\),
this can add appreciable bias to the resulting estimates (e.g. Fan and
Li 2001). One method proposed for overcoming this limitation is the
adaptive lasso (Zou 2006). Instead of penalizing parameters directly,
each parameter is scaled by the un-penalized estimated (MLE parameter
estimates in SEM). The adaptive lasso results in: \[
F_{alasso} = F_{ML} + \lambda \| \theta_{ML}^{-1} * \theta_{pen} \|_{1}
\] \noindent
with, following the same form for the lasso, the soft-thresholding
update is:

\[
S(\theta_{j})= sign(\theta_{j})(|\theta_{j}|-\frac{\lambda}{2|\theta_{j}|})_{+}
\] \noindent
In this, larger penalties are given for non-significant (smaller)
parameters, limiting the bias in estimating larger, significant
parameters. Note that one limitation of this approach for SEM models is
that the model needs to be estimable with MLE. Particularly for models
with large numbers of variables, in relation to sample size, this may
not be possible.

@@

Two additional penalties that overcome some of the deficiencies of the
lasso, producing sparser solutions, include the smoothly clipped
absolute deviation penalty (SCAD; Fan and Li 2001) and the minimax
concave penalty (MCP; Zhang 2010). In comparison to the lasso, both the
SCAD and MCP have much smaller penalties for large parameters, where the
amount of penalty for small penalties is similar to the lasso, as is
evident in Figure 2.

@@

\noindent
for \(\gamma > 2\). As the the penalty in equation 11 is non-convex (as
is the MCP), this makes computation more difficult. However, in the
context of SEM this can be seen as less problematic, as equation 3 is
also non-convex.

@@

\noindent
for \(\gamma > 0\). As seen in Figure 2, this results in similar amount
of shrinkage for smaller estimates in comparison to the SCAD, however,
less for larger estimates. For both the SCAD and MCP, both the
\(\gamma\) and \(\lambda\) parameters are used as hyper-parameters. This
involves testing models over a two-dimensional array of parameters,
however, \(\gamma\) can be fixed to 3.7 per Fan and Li (2001).

@@

RegSEM is implemented as the \pkg{regsem} package (Jacobucci 2017) in
the \proglang{R} statistical environment (R Core Team 2017). To estimate
the maximum likelihood fit of the model, \pkg{regsem} uses
\textit{Reticular Action Model} (RAM; J. J. McArdle and McDonald 1984,
McArdle (2005)) notation to derive an implied covariance matrix. The
parameters of each SEM are translated into three matrices: the
\textit{filter} (\textit{F}), the \textit{asymmetric} (\textit{A};
directed paths; e.g.~factor loadings or regressions), and the
\textit{symmetric} (\textit{S}; undirected paths; e.g.~covariances or
variances). See Jacobucci, Grimm, and McArdle (2016) for more detail on
RAM notation.

Syntax for using the \pkg{regsem} is based on the \pkg{lavaan} package
(Rosseel 2012) for structural equation models. \pkg{lavaan} is a general
SEM software program that can fit a wide array of models with various
estimation methods. To use \pkg{regsem}, the user has to first fit the
model in \pkg{lavaan}. As an example, below is the code for a one factor
model fit in lavaan with the Holzinger and Swineford (1939) dataset.

@@

The \code{out.reg} object contains two components, \code{out.reg[[1]]}
has the parameter estimates for each of the 15 models,

\begin{CodeChunk}
\begin{CodeInput}
head(round(out.reg[[1]],2),5)
\end{CodeInput}
\begin{CodeOutput}
     f1 -> x1 f1 -> x2 f1 -> x3 f1 -> x4 f1 -> x5 f1 -> x6 f1 -> x7
[1,]     0.51     0.26     0.25     0.98     1.08     0.92     0.20
[2,]     0.41     0.17     0.17     0.89     0.98     0.83     0.12
[3,]     0.33     0.10     0.10     0.83     0.91     0.77     0.06
[4,]     0.27     0.04     0.04     0.78     0.87     0.72     0.01
[5,]     0.21     0.00     0.00     0.75     0.83     0.69     0.00
     f1 -> x8 f1 -> x9 x1 ~~ x1 x2 ~~ x2 x3 ~~ x3 x4 ~~ x4 x5 ~~ x5
[1,]     0.20     0.31     1.10     1.31     1.21     0.38     0.49
[2,]     0.13     0.23     1.11     1.32     1.22     0.38     0.48
[3,]     0.08     0.17     1.13     1.33     1.23     0.38     0.47
[4,]     0.03     0.12     1.15     1.35     1.25     0.38     0.46
[5,]     0.00     0.07     1.18     1.36     1.26     0.39     0.46
     x6 ~~ x6 x7 ~~ x7 x8 ~~ x8 x9 ~~ x9
[1,]     0.36     1.15     0.98     0.92
[2,]     0.36     1.15     0.99     0.93
[3,]     0.36     1.16     1.00     0.94
[4,]     0.37     1.17     1.01     0.96
[5,]     0.38     1.18     1.02     0.97
\end{CodeOutput}
\end{CodeChunk}

\noindent
while \code{out.reg[[2]]} looks like:

\begin{CodeChunk}
\begin{CodeInput}
round(out.reg[[2]],2)
\end{CodeInput}
\begin{CodeOutput}
      lambda conv rmsea     BIC
 [1,]   0.00    0  0.19 7805.18
 [2,]   0.05    0  0.19 7815.39
 [3,]   0.10    0  0.20 7840.59
 [4,]   0.15    0  0.21 7875.84
 [5,]   0.20    0  0.20 7886.15
 [6,]   0.25    0  0.21 7911.82
 [7,]   0.30    0  0.21 7934.45
 [8,]   0.35    0  0.22 7960.67
 [9,]   0.40    0  0.23 7991.03
[10,]   0.45    0  0.23 8009.42
[11,]   0.50    0  0.23 8034.81
[12,]   0.55    0  0.24 8069.49
[13,]   0.60    0  0.29 8360.54
[14,]   0.65    0  0.29 8360.53
[15,]   0.70    0  0.29 8360.52
\end{CodeOutput}
\end{CodeChunk}

In this, the user can examine the penalty (lambda), whether the model
converged (``conv''=0), and the fit of each model. By default, two fit
indices are output, both the root mean square error of approximation
(RMSEA; Steiger and Lind 1980), and the Bayesian information criteria
(BIC; Schwarz 1978). Both the RMSEA and BIC take into account the
degrees of freedom of the model, an important point for model selection
in the presence of lasso penalties (and other penalties that set
parameters to zero). Based on work done on degrees of freedom for lasso
regression, as Zou, Hastie, and Tibshirani (2007) proved that the number
of nonzero coefficients is an unbiased estimate of the degrees of
freedom for regression. As the penalty increases, select parameter are
set to zero, thus the degrees of freedom increases, which for fit
indices that include the degrees of freedom in the calculation, means
that although the likelihood of the model may only get worse (increase),
both the RMSEA and BIC can improve (decrease) due to the impact of
increased degrees of freedom.

Instead of examining the \code{out.reg[[1]]} output matrix of parameter
estimates, users also have the option to plot the trajectory of each of
the penalized parameters. This is accomplished with
\code{plot_cv(out.reg,pars=1:9)}, resulting in:

@@

\begin{center}\includegraphics{draft1_files/figure-latex/unnamed-chunk-6-1} \end{center}

@@

After a final model (penalty) is chosen, users have the option either
just use the output from \code{cv_regsem()}, or the final model can be
re-run with either \code{regsem()} or \code{multi_optim()} to attain
additional information. In the model above, the best fitting penalty,
according to the BIC, is \(\lambda=0\). However, for demonstration
purposes, we can choose a penalty of 0.2. With this, the model is re-run
with \code{regsem()}

@@

For the purposes of comparison, each of the 20 covariate regressions
were penalized using the lasso, adaptive lasso, SCAD, and MCP, and
compared to the maximum likelihood estimates. In this model, the data
were simulated to have two large effects (both \textit{c1} parameters),
two small effects (both \textit{c2} parameters) and sixteen true zero
effects (\textit{c3-c10} parameters). Note that the covariates were
simulated to have zero covariance among each variable. If there was
substantial collinearity among covariates, the elastic net would be more
appropriate to simultaneously select predictors while also accounting
for the collinearity. The parameter estimates corresponding the the best
fit of the BIC are has the fit of each model, resulting in Table 1,
created using the \pkg{xtable} package (Dahl 2009).

@@

While every regularization method erroneously set both simulated true
intercept effects as zero (as in maximum likelihood, which accounts for
these errors), both the adaptive lasso and SCAD correctly identified
every true zero effect. The lasso identified two false effects while the
MCP mistakenly identified one. Additionally, the lasso estimation of the
true effects was attentuated in comparison to the other regularization
methods. This is in line with previous research (Fan and Li 2001),
necessitating the use of a two-step relaxed lasso method (Meinshausen
2007, see Jacobucci, Grimm, and McArdle (2016)) As expected given the
small ratio between number of estimated parameters and sample size,
maximum likelihood mistakenly identified 3 false effects as significant.

To compare the performance of each penalization method further,
particularly in the presence of small number of parameters to sample
size ratios, a small scale simulation study was conducted. The same
model and effects was kept, but the sample size was varied to include
80, 200, and 1000 to demonstrate how maximum likelihood estimation of
effects improves as sample size increases, while each of the
regularization methods performs well regardless of sample size. Each run
was replicated 200 times. For each regularization method, the BIC was
used to choose a final model among the 40 penalty vales. The results are
displayed in Table 2.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & N & ML & lasso & alasso & SCAD & MCP \\ 
  \hline
1 & 80.00 & 0.08 & 0.08 & \textbf{0.04} & 0.05 & \textbf{0.04} \\ 
  2 & 200.00 & 0.06 & 0.05 & \textbf{0.02} & 0.03 & 0.03 \\ 
  3 & 1000.00 & 0.05 & 0.08 & \textbf{0.01} & \textbf{0.01} & 0.28 \\ 
  4 & 80.00 & 0.33 & \textbf{0.31} & 0.35 & 0.35 & 0.45 \\ 
  5 & 200.00 & \textbf{0.19} & \textbf{0.19} & 0.23 & 0.26 & 0.45 \\ 
  6 & 1000.00 & \textbf{0.00} & \textbf{0.00} & \textbf{0.00} & 0.01 & 0.14 \\ 
   \hline
\end{tabular}
    \caption{Results from the simulation using the model in Figure 1. Each condition was replicated 200 times. False positives represent concluding that the simulated regressions of zero were concluded as nonzero. False negatives are concluding that either the simulated regression values of 1 or 0.2 are in fact zero. Bolded values represent the smallest error per condition.}
\end{table}

For false positive, the adaptive lasso demonstrated the best
performance, where the performance of maximum likelihood estimation
leveled off at the 0.05 level at a sample size of 1000 as expected. For
false negatives, lasso penalties demonstrated similar results to maximum
likelihood. This was expected given the tendency of the lasso to
under-penalize small coefficients in comparison to the other
regularization methods. The adaptive lasso and SCAD demonstrated
slightly worse results, however, outside of the MCP, each method made
either zero or near zero errors at a sample size of 1000. The poor
performance of the MCP may be in part due to fixing the \(\gamma\)
penalty to 3.7. Varying this parameter may improve the performance of
the method. In summary, the regularization methods demonstrated an
improvement over maximum likelihood, particularly at small samples, for
a model that had a large number of estimated parameters.
